Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case

### 2019-01-23 ICML-submission/appendix-nearest-neighbor-algorithm.tex
Fixes
Clean-ups

\begin{algorithm}[H]
\caption{\textsc{Nearest-Neighbor Algorithm}}
\begin{algorithmic}[1]
\REQUIRE{Number of classes $K$, number of rounds $T$.}
\REQUIRE{Inner product space $(V,\ip{\cdot}{\cdot})$.}
\STATE $S \gets \emptyset$.
\FOR{$t=1,2,\ldots,T$:}
\IF{$\min_{(x,y) \in S} \norm{x_t - x} \le \gamma$}
  \STATE{Find nearest neighbor \\ $(\widetilde{x}, \widetilde{y}) = \argmin_{(x,y) \in S} \norm{x_t - x}$}
	\STATE{Predict $\widehat{y}_t = \widetilde{y}$}
\ELSE
	\STATE{Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$}
	\STATE{Receive feedback $z_t = \indicator{\widehat{y}_t \neq y_t}$}
  \label{line:nearest-neighbor-explore}
	\IF{$z_t = 0$}
    \STATE $S \gets S \cup \cbr{(x_t, \widehat{y}_t)}$.
  \ENDIF
\ENDIF
\ENDFOR
\end{algorithmic}
\label{algorithm:nearest-neighbor}
\end{algorithm}

In this section we analyze \textsc{Nearest-Neighbor Algorithm} shown as
Algorithm~\ref{algorithm:nearest-neighbor}. The algorithm is based on the idea
that two examples that are close to each other must have the same label. Lemma
below formalizes this intuition.

\begin{lemma}
\label{lemma:nearest-neighbor}
Let $(V,\ip{\cdot}{\cdot})$ be a vector space, $K$ be a positive integer and let
$\gamma$ be a positive real number. Suppose $(x_1,y_1), (x_2,y_2), \dots, (x_T,
y_T) \in V \times \{1,2,\dots,K\}$ are labeled examples that are weakly linearly
separable with margin $\gamma$. For $i$, $j$ in $\cbr{1,2,\dots,T}$, if
$\norm{x_i - x_j}_2 \le \gamma$ then $y_i = y_j$.
\end{lemma}

@@

\begin{theorem}
\label{theorem:margin_at_upper}
Let $K$ and $d$ be positive integers and let $\gamma$ be a positive real number.
Suppose $(x_1,y_1), \ldots, (x_T, y_T) \in \R^d \times \{1,2,\dots,K\}$ are
labeled examples that are weakly linearly separable with margin $\gamma$ and
satisfy $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$. Then, the expected
number of mistakes made by Algorithm~\ref{algorithm:nearest-neighbor} is at most
$$
(K-1) \left( \frac{R}{\gamma} \right)^d \; .
$$
\end{theorem}

\begin{proof}
Let $M$ be the number of mistakes made by the algorithm. Let $b_t$ be the
indicator that line~\ref{line:nearest-neighbor-explore} is executed at timestep
$t$, i.e. we fall into the ``else'' case. Note that if $b_t = 0$, then by
Lemma~\ref{lemma:nearest-neighbor}, the prediction $\widehat{y}_t$ must equal
$y_t$, i.e. $z_t = 0$. Therefore, $M = \sum_{t=1}^T z_t = \sum_{t=1}^T b_t z_t$.
Let $U = \sum_{t=1}^T b_t (1-z_t)$. Clearly, $|S| = U$. Since $S$ is a
$\gamma$-packing over $\cbr{x \in \R^d ~:~ \norm{x} \le R}$, the size of $S$ is
at most $(\frac R \gamma)^d$, which is also an upper bound on $U$.

Note that when $b_t = 1$, $\widehat{y}_t$ is chosen uniformly at random, we have
$$
\Exp[ z_t | b_t = 1] = \frac{K-1}{K} \; .
$$
Therefore,
$$
\Exp[M] = \Exp \left[ \sum_{t=1}^T b_t z_t \right] = \frac{K-1}{K} \Exp \left[ \sum_{t=1}^T b_t \right] \; .
$$
On the other hand,
$$
\Exp[U] = \Exp \left[ \sum_{t=1}^T b_t (1-z_t) \right] = \frac 1 K \Exp \left[\sum_{t=1}^T b_t \right] \; .
$$
Therefore,
$$
\Exp[M] = (K-1) \Exp[U] \leq (K-1) \left(\frac{R}{\gamma} \right)^d \; .
$$
\end{proof}

>>>>

\begin{algorithm}[H]
\SetAlgoLined
\LinesNumbered
\caption{\textsc{Nearest-Neighbor Algorithm}
\label{algorithm:nearest-neighbor}
}
\textbf{Require:} Number of classes $K$, number of rounds $T$. \\
\textbf{Require:} Inner product space $(V,\ip{\cdot}{\cdot})$. \\
\nl Initialize $S \gets \emptyset$ \\
\nl \For{$t=1,2,\ldots,T$:}{
\nl  \If{$\min_{(x,y) \in S} \norm{x_t - x} \le \gamma$}{
\nl    Find nearest neighbor \\
       \qquad \qquad $(\widetilde{x}, \widetilde{y}) = \argmin_{(x,y) \in S} \norm{x_t - x}$ \\
\nl	  Predict $\widehat{y}_t = \widetilde{y}$
  }
  \Else{
\nl	  Predict $\widehat y_t \sim \text{Uniform}(\{1,2,\dots,K\})$
      \label{line:nearest-neighbor-explore} \\
\nl	  Receive feedback $z_t = \indicator{\widehat{y}_t \neq y_t}$ \\
\nl	  \If{$z_t = 0$}{
\nl      $S \gets S \cup \cbr{(x_t, \widehat{y}_t)}$
    }
  }
}
\end{algorithm}

In this section we analyze \textsc{Nearest-Neighbor Algorithm} shown as
Algorithm~\ref{algorithm:nearest-neighbor}. The algorithm is based on obvious
the idea that under weak linear assumption two examples that are close to each
other must have the same label. Lemma below formalizes this intuition.

\begin{lemma}[Separation lemma]
\label{lemma:separation-lemma}
Let $(V,\ip{\cdot}{\cdot})$ be a vector space, $K$ be a positive integer and let
$\gamma$ be a positive real number. Suppose $(x_1,y_1), (x_2,y_2), \dots, (x_T,
y_T) \in V \times \{1,2,\dots,K\}$ are labeled examples that are weakly linearly
separable with margin $\gamma$. For $i$, $j$ in $\cbr{1,2,\dots,T}$, if
$\norm{x_i - x_j}_2 \le \gamma$ then $y_i = y_j$.
\end{lemma}

@@

We also need to define several notions. A subset $S \subseteq \R^d$ is called a
$\gamma$-packing if for any $x,x' \in S$ such that $x \neq x'$ we have $\norm{x -
x'} > \gamma$. The following lemma is standard. Also recall that $B(x,R) = \{ x'
\in \R^d ~:~ \norm{x' - x} \le R \}$ denotes the closed ball of radius $R$
centered a point $x$.

\begin{lemma}[Size of $\gamma$-packing]
\label{lemma:size-of-packing}
Let $\gamma$ and $R$ be positive real numbers.
If $S \subseteq B(\zero,R) \subseteq \R^d$ is a $\gamma$-packing then
$$
|S| \le \left( \frac{2R}{\gamma} + 1 \right)^d \; .
$$
\end{lemma}

\begin{proof}
If $S$ is a $\gamma$-packing then $\{ B(x,\gamma/2) ~:~ x \in S \}$
is a collection of disjoint balls of radius $\gamma$ that fit into $B(\zero,R + \gamma/2)$.
Thus,
$$
|S| \cdot \Vol(B(\zero, \gamma/2)) \le \Vol(B(\zero,R + \gamma/2))
$$
Hence,
\begin{align*}
|S|
& \le \frac{\Vol(B(\zero,R + \gamma/2))}{\Vol(B(\zero, \gamma/2))} \\
& = \left( \frac{R + \gamma/2}{\gamma/2} \right)^d \\
& = \left( \frac{2R}{\gamma} + 1 \right)^d \; .
\end{align*}
\end{proof}

\begin{theorem}[Mistake upper bound for \textsc{Nearest-Neighbor Algorithm}]
\label{theorem:mistake-bound-for-nearest-neighbor-algorithm}
Let $K$ and $d$ be positive integers and let $\gamma,R$ be a positive real
numbers. Suppose $(x_1,y_1), \ldots, (x_T, y_T) \in
\R^d \times \{1,2,\dots,K\}$ are labeled examples that are weakly linearly
separable with margin $\gamma$ and satisfy $\norm{x_1}, \norm{x_2}, \dots,
\norm{x_T} \le R$. Then, the expected number of mistakes made by
Algorithm~\ref{algorithm:nearest-neighbor} is at most
$$
(K-1) \left( \frac{2R}{\gamma} + 1\right)^d \; .
$$
\end{theorem}

\begin{proof}
Let $M$ be the number of mistakes made by the algorithm. Let $b_t$ be the
indicator that line~\ref{line:nearest-neighbor-explore} is executed at timestep
$t$, i.e. we fall into the ``else'' case. Note that if $b_t = 0$, then by
Lemma~\ref{lemma:separation-lemma}, the prediction $\widehat{y}_t$ must equal
$y_t$, i.e. $z_t = 0$. Therefore, $M = \sum_{t=1}^T z_t = \sum_{t=1}^T b_t z_t$.
Let $U = \sum_{t=1}^T b_t (1-z_t)$. Clearly, $|S| = U$. Since $S \subseteq B(\zero, R)$
is a $\gamma$-packing, $U = |S| \le (\frac{2R}{\gamma} + 1)^d$.

Note that when $b_t = 1$, $\widehat{y}_t$ is chosen uniformly at random, we have
$$
\Exp[ z_t ~|~ b_t = 1] = \frac{K-1}{K} \; .
$$
Therefore,
$$
\Exp[M] = \Exp \left[ \sum_{t=1}^T b_t z_t \right] = \frac{K-1}{K} \Exp \left[ \sum_{t=1}^T b_t \right] \; .
$$
On the other hand,
$$
\Exp[U] = \Exp \left[ \sum_{t=1}^T b_t (1-z_t) \right] = \frac 1 K \Exp \left[\sum_{t=1}^T b_t \right] \; .
$$
Therefore,
$$
\Exp[M] = (K-1) \Exp[U] \leq (K-1) \left(\frac{2R}{\gamma} + 1 \right)^d \; .
$$
\end{proof}

### 2019-01-23 ICML-submission/mistake-lower-bound-for-ignorant-algorithms.tex
Clean-ups

The ignorant algorithm is formally defined as follows.
\begin{definition} \label{definition:ignorant-algorithm}
A bandit classification algorithm $\mathcal{A}$ is ignorant if for all
$t\in[T]$, $\widehat y_t\in \{1,2\}$ is drawn from some distribution that is
jointly determined by
\begin{enumerate}
\item $x_t$
\item $H_t \triangleq \left( (x_{\tau_1}, y_{\tau_1}), (x_{\tau_2}, y_{\tau_2}), \ldots, (x_{\tau_{n_t}}, y_{\tau_{n_t}}) \right)$,
where $1 \le \tau_1 < \dots < \tau_{n_t} < t$, and $\{\tau_1, \ldots, \tau_{n_t}\}$
are the rounds in which the algorithm predicts correctly (that is,
$\{\tau_1, \ldots, \tau_{n_t}\} = \{1 \le \tau < t ~:~ \widehat y_\tau=y_\tau\}$).
\end{enumerate}
We use $p_t(\cdot|x_t)\in \{p\in \R_+^2 ~:~ p^\top\mathbf{1}=1\}$ to denote
the distribution $\widehat y_t$ is drawn from. Note that we let the current
feature vector $x_t$ be an argument to $p_t$ to incorporate its dependency on
$x_t$. Therefore, for an ignorant algorithm, the function $p_t(\cdot|\cdot)$ is
only determined by $H_t$.
\end{definition}

@@

\begin{theorem}[Mistake Lower Bound for Ignorant Algorithms]
\label{theorem:ignorant_lower_bound}
Suppose $\mathcal{A}$ is an ignorant bandit classification algorithm. There
exists an adversary that can sequentially choose $(x_1, y_1), (x_2, y_2),
\ldots, (x_T, y_T) \in \R^d\times \{1,2\}$, such that $\forall t\in[T],
\|x_t\|\le 1$, $\{(x_t, y_t)\}_{t=1}^T$ is $\gamma$-strongly linearly
separable, and the expected number of mistakes made by $\mathcal{A}$ is lower
bounded as
\begin{align*}
\Exp\left[ \sum_{t=1}^T \indicator{\widehat y_t \neq y_t} \right]
\ge \frac{1}{10}\min\left\{\sqrt{T}, \left(\frac{1}{160\gamma}\right)^{\frac{d-2}{4}} \right\} \; .
\end{align*}
\end{theorem}

@@

\begin{lemma}
\label{lemma:embed_d_gamma}
Let $\gamma \in (0,\frac 1 {160})$, let $d$ be a positive integer and let $N = (\Omega(\frac{1}{\gamma}))^{(d-2)/2}$.
There exist vectors $v_1, v_2, \dots, v_N, u_1, u_2, \dots, u_N \in \R^d$ such that for all $i, j \in \{1,2,\dots,N\}$
\begin{align*}
\norm{v_i} & \le 1 \, \\
\norm{u_i} & \le 1 \, \\
\ip{u_i}{v_j} & \ge \gamma & & \text{if $i=j$,} \\
\ip{u_i}{v_j} & \le -\gamma & & \text{if $i \neq j$.}
\end{align*}
\end{lemma}

\begin{proof}
By Lemma 6 of~\cite{Long-1995}, there exists a set of $N =
(\frac{1}{2\sqrt{40\gamma}})^{d-2}$ $(d-1)$-dimensional vectors $z_1, \ldots,
z_N$ on the unit sphere $\mathbb{S}^{d-2}$, such that $\theta(z_i, z_j) \ge
\sqrt{40 \gamma}$. As $\cos\theta \le 1-\theta^2/5$ for every $\theta \in
[0,\pi]$, this implies that
$$
\ip{z_i}{z_j} \begin{cases} = 1 ,& i = j, \\ \le 1- 8\gamma ,& i \neq j. \end{cases}
$$

Define $v_i = (\frac12 z_i, \frac12)$, and $u_i = (\frac{1}{2} z_i, -\frac{1}{2}(1-4\gamma))$ for all $i \in \{1,2,\dots,N\}$.
It can be easily checked that for all $i$,
$\norm{v_i} \le 1$ and $\norm{u_i} \le 1$. In addition,
$$
\ip{u_i}{v_j}
= \frac{1}{4} \ip{z_i}{z_j} - \frac {1-4\gamma} 4 \begin{cases} \ge \gamma ,& i = j, \\ \le -\gamma ,& i \neq j. \end{cases}
$$
\end{proof}

\begin{proof}[Proof for Theorem~\ref{theorem:ignorant_lower_bound} (PartI: Mistake lower bound)]
We consider the strategy for the adversary described in Algorithm~\ref{algorithm:adversary-strategy}.
\begin{algorithm}[h]
\caption{\textsc{Adversary's strategy}}
\label{algorithm:adversary-strategy}
\begin{algorithmic}[1]
{
\STATE{\textbf{Define} $q_0=\max\left\{\frac{1}{\sqrt{T}}, \frac{1}{\sqrt{N}}\right\}$, where $N$ is defined in Lemma~\ref{lemma:embed_d_gamma}. Also, let $v_1, \ldots, v_N$ be those defined in Lemma~\ref{lemma:embed_d_gamma}.}
\STATE{\textbf{Initialize} $\textsc{phase}= 1$. }
\FOR{$t=1,\dots,\min\{T,N\}$}
    \IF{$\textsc{phase}=1$}
       \IF{$p_t(1|v_t)\ge 1-q_0$}
          \STATE{$(x_t, y_t)\leftarrow (v_t, 1)$}
       \ELSE
          \STATE{$(x_t, y_t)\leftarrow (v_t, 2)$}
          \STATE $\textsc{phase}\leftarrow 2$
       \ENDIF
    \ELSIF{$\textsc{phase}=2$}
       \STATE $(x_t, y_t)\leftarrow (x_{t-1}, y_{t-1})$.
    \ENDIF
\ENDFOR
\FOR{$t=\min\{T,N\}+1, \ldots, T$}
    \STATE $(x_t, y_t)\leftarrow (x_{t-1}, y_{t-1})$.
\ENDFOR

>>>>

The ignorant algorithm is formally defined as follows.
\begin{definition}
\label{definition:ignorant-algorithm}
A bandit classification algorithm $\calA$ is ignorant if for all $t \in
\{1,2,\dots,T\}$, $\widehat y_t\in \{1,2\}$ is drawn from some distribution that
is jointly determined by
\begin{enumerate}
\item $x_t$
\item $H_t \triangleq \left( (x_{\tau_1}, y_{\tau_1}), (x_{\tau_2}, y_{\tau_2}), \ldots, (x_{\tau_{n_t}}, y_{\tau_{n_t}}) \right)$,
where $1 \le \tau_1 < \dots < \tau_{n_t} < t$, and $\{\tau_1, \ldots, \tau_{n_t}\}$
are the rounds in which the algorithm predicts correctly (that is,
$\{\tau_1, \ldots, \tau_{n_t}\} = \{1 \le \tau < t ~:~ \widehat y_\tau=y_\tau\}$).
\end{enumerate}
We use $p_t(\cdot|x_t)\in \{p\in \R_+^2 ~:~ p^\top\mathbf{1}=1\}$ to denote
the distribution $\widehat y_t$ is drawn from. Note that we let the current
feature vector $x_t$ be an argument to $p_t$ to incorporate its dependency on
$x_t$. Therefore, for an ignorant algorithm, the function $p_t(\cdot|\cdot)$ is
only determined by $H_t$.
\end{definition}

@@

\begin{theorem}[Mistake Lower Bound for Ignorant Algorithms]
\label{theorem:ignorant_lower_bound}
Suppose $\calA$ is an ignorant bandit classification algorithm. There
exists an adversary that can sequentially choose $(x_1, y_1), (x_2, y_2), \dots,
(x_T, y_T) \in \R^d\times \{1,2\}$ so that the examples $\gamma$-strongly
linearly separable, $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$, and the
expected number of mistakes made by $\calA$ is at least
$$
\frac{1}{10}\min\left\{\sqrt{T}, \left(\frac{1}{160\gamma}\right)^{\frac{d-2}{4}} \right\} \; .
$$
\end{theorem}

@@

\begin{lemma}
\label{lemma:embed_d_gamma}
Let $\gamma \in (0,\frac 1 {160})$, let $d$ be a positive integer and let $N = (\Omega(\frac{1}{\gamma}))^{(d-2)/2}$.
There exist vectors $u_1, u_2, \dots, u_N, v_1, v_2, \dots, v_N \in \R^d$ such that for all $i, j \in \{1,2,\dots,N\}$,
\begin{align*}
\norm{u_i} & \le 1 \; , \\
\norm{v_j} & \le 1 \; , \\
\ip{u_i}{v_j} & \ge \gamma & & \text{if $i=j$,} \\
\ip{u_i}{v_j} & \le -\gamma & & \text{if $i \neq j$.}
\end{align*}
\end{lemma}

\begin{proof}
By Lemma 6 of~\cite{Long-1995}, there exists a set of $N =
(\frac{1}{2\sqrt{40\gamma}})^{d-2}$ $(d-1)$-dimensional vectors $z_1, \ldots,
z_N$ on the unit sphere $\mathbb{S}^{d-2}$, such that $\theta(z_i, z_j) \ge
\sqrt{40 \gamma}$. As $\cos\theta \le 1-\theta^2/5$ for every $\theta \in
[0,\pi]$, this implies that
\begin{align*}
\ip{z_i}{z_j} &= 1 && \text{if $i = j$,} \\
\ip{z_i}{z_j} &\le 1 - 8\gamma && \text{if $i \neq j$.}
\end{align*}

Define $v_i = (\frac12 z_i, \frac12)$, and $u_i = (\frac{1}{2} z_i, -\frac{1}{2}(1-4\gamma))$ for all $i \in \{1,2,\dots,N\}$.
It can be easily checked that for all $i$,
$\norm{v_i} \le 1$ and $\norm{u_i} \le 1$. Additionally,
$$
\ip{u_i}{v_j} = \frac{1}{4} \ip{z_i}{z_j} - \frac {1-4\gamma} 4 \; .
$$
Thus,
\begin{align*}
\ip{u_i}{v_j} &\ge \gamma && \text{if $i=j$,} \\
\ip{u_i}{v_j} &\le -\gamma && \text{if $i \neq j$.}
\end{align*}
\end{proof}

\begin{proof}[Proof for Theorem~\ref{theorem:ignorant_lower_bound} (Part I: Mistake lower bound)]
We consider the strategy for the adversary described in Algorithm~\ref{algorithm:adversary-strategy}.
\begin{algorithm}[h]
\caption{\textsc{Adversary's strategy}}
\label{algorithm:adversary-strategy}
\begin{algorithmic}[1]
{
\STATE{\textbf{Define} $q_0=\max\left\{\frac{1}{\sqrt{T}}, \frac{1}{\sqrt{N}}\right\}$, where $N$ is defined in Lemma~\ref{lemma:embed_d_gamma}. Also, let $v_1, \ldots, v_N$ be those defined in Lemma~\ref{lemma:embed_d_gamma}.}
\STATE{\textbf{Initialize} $\textsc{phase}= 1$. }
\FOR{$t=1,\dots,\min\{T,N\}$}
    \IF{$\textsc{phase}=1$}
       \IF{$p_t(1|v_t)\ge 1-q_0$}
          \STATE{$(x_t, y_t)\leftarrow (v_t, 1)$}
       \ELSE
          \STATE{$(x_t, y_t)\leftarrow (v_t, 2)$}
          \STATE $\textsc{phase}\leftarrow 2$
       \ENDIF
    \ELSIF{$\textsc{phase}=2$}
       \STATE $(x_t, y_t)\leftarrow (x_{t-1}, y_{t-1})$.
    \ENDIF
\ENDFOR
\FOR{$t=\min\{T,N\}+1, \ldots, T$}
    \STATE $(x_t, y_t)\leftarrow (x_{t-1}, y_{t-1})$.
\ENDFOR

### 2019-01-23 ICML-submission/np-hardness-of-the-weak-labeling-problem.tex
Fixes
Clean-ups

\section{NP-hardness of the weak labeling problem}

@@

We use for $1,2,\dots,K$ for strong labels and $\overline{1}, \overline{2},
\dots, \overline{K}$ for weak labels. We adopt the convention that
$\overline{\overline{i}} = i$ for any positive integer $i$. Formally, the weak
labeling problem can be described as below:
\begin{figure}[H]
\begin{framed}
\begin{center}
    \textbf{Weak Labeling}
\end{center}
\textbf{Given:} Feature-label pairs $(x_1, y_1)$, $(x_2, y_2)$, \dots, $(x_T, y_T)$ in $\{0,1\}^d \times \{1,2,\dots, K, \overline{1}, \overline{2}, \dots, \overline{K}\}$. \\
\textbf{Question:} Do there exist $ w_1, w_2, \dots, w_K \in \R^d$ such that for all $t=1,2,\dots,T$,
\begin{align*}
& y_t \in \{1,2,\dots,K\} \Longrightarrow  \\
& \quad \forall i \in \{1,2,\dots,K\} \setminus \{y_t\} \quad \ip{w_{y_t}}{x_t}  > \ip{w_i}{x_t} \; , \\
& \text{and} \\
& y_t \in \{\overline{1}, \overline{2},\dots, \overline{K}\} \Longrightarrow \\
& \quad \exists i \in \{1,2,\dots,K\} \quad \ip{w_i}{x_t} > \ip{w_{\overline{y_t}}}{x_t} \; ?
\end{align*}
\end{framed}
\end{figure}

@@

\begin{proof}
$(\Longrightarrow)$ Let $S_1, S_2$ be the solution of the set splitting problem. Define
$$
w_1 = \left(a_1, a_2, \cdots, a_N, -\frac{1}{2}\right),
$$
where for all $i \in \{1,2,\dots,N\}$, $a_i=1$ if $i\in S_1$ and $a_i=-N$ if
$i\notin S_1$. Similarly, define
$$
w_2 = \left(b_1, b_2, \cdots, b_N, -\frac{1}{2}\right),
$$
where for all $i \in \{1,2,\dots,N\}$, $b_i=1$ if $i \in S_2$ and $b_i=-N$ if
$i\notin S_2$. Finally, define
$$
w_3 = (0,0,\cdots, 0),
$$
the zero vector. To see this is a solution for the weak labeling problem, we
verify separately for Type 1-3 samples defined above. For Type 1 sample, we have
$$
\ip{w_3}{x} = 0 > -\frac{1}{2} = \ip{w_1}{x}=\ip{w_2}{x}.
$$
For a Type 2 sample that corresponds to index $i$, we have either $i\in S_1$ or
$i\in S_2$ because $S_1\cup S_2 = \{1,2,\dots,N\}$ is guaranteed. Thus, either
$a_i=1$ or $b_i=1$. If $a_i=1$ is the case, then
$$
\ip{w_1}{x} = a_i - \frac{1}{2} = \frac{1}{2} > 0 = \ip{w_3}{x};
$$
similarly if $b_i=1$, we have $\ip{w_2}{x}>\ip{w_3}{x}$. \\ For a Type 3 sample
that corresponds to index $j$, Since $c_j \not\subset S_1$, there exists some
$i'\in c_j$ and $i'\notin S_1$. Thus we have $x_{i'}=1$, $a_{i'}=-N$, and
therefore
\begin{align*}
\ip{w_1}{x}
& = a_{i'}x_{i'} + \sum_{i\in \{1,2,\dots,N\} \setminus \{i'\}} a_ix_i - \frac{1}{2} \\
& \le -N + (N-1)-\frac{1}{2} \\
& < 0 = \ip{w_3}{x} \; .
\end{align*}
Because $c_j \not\subset S_2$ also holds, we also have
$\ip{w_2}{x}<\ip{w_3}{x}$. This direction is therefore proved. \\
\ \\
$(\Longleftarrow)$ Given the solution $w_1, w_2, w_3$ of the weak labeling problem, we define
\begin{align*}
S_1 &= \left\{i \in \{1,2,\dots,n\} ~:~ \ip{w_1-w_3}{(\e_i, 1)} > 0 \right\}, \\
S_2 &= \left\{i \in \{1,2,\dots,n\} ~:~ \ip{w_2-w_3}{(\e_i, 1)} > 0 \text{\ and\ } i\notin S_1 \right\}.
\end{align*}
It is not hard to see $S_1 \cap S_2 = \emptyset$ and $S_1\cup S_2 =
\{1,2,\dots,N\}$. The former is because $S_2$ only includes elements that are
not in $S_1$. For the latter, note that $(\e_i, 1)$ is the feature vector for
Type 2 samples. Because Type 2 samples all have label $\overline{3}$, for any $i
\in \{1,2,\dots,N\}$, one of the following must hold: $\ip{w_1-w_3}{(\e_i,
1)}>0$ or $\ip{w_2-w_3}{(\e_i, 1)}>0$. This implies $i\in S_1$ or $i\in S_2$.

>>>>

\section{NP-hardness of the weak labeling problem}
\label{section:np-hardness-of-weak-labeling-problem}

@@

We use symbols $1,2,\dots,K$ for strong labels and symbols $\overline{1},
\overline{2}, \dots, \overline{K}$ for weak labels. We adopt the convention that
$\overline{\overline{i}} = i$ for any symbol $i$. Formally, the weak
labeling problem can be described as below:
\begin{figure}[H]
\begin{framed}
\begin{center}
    \textbf{Weak Labeling}
\end{center}
\textbf{Given:} Feature-label pairs $(x_1, y_1)$, $(x_2, y_2)$, \dots, $(x_T, y_T)$ in $\{0,1\}^d \times \{1,2,\dots, K, \overline{1}, \overline{2}, \dots, \overline{K}\}$. \\
\textbf{Question:} Do there exist $ w_1, w_2, \dots, w_K \in \R^d$ such that for all $t=1,2,\dots,T$,
\begin{align*}
& y_t \in \{1,2,\dots,K\} \Longrightarrow  \\
& \quad \forall i \in \{1,2,\dots,K\} \setminus \{y_t\} \quad \ip{w_{y_t}}{x_t}  > \ip{w_i}{x_t} \; , \\
& \text{and} \\
& y_t \in \{\overline{1}, \overline{2},\dots, \overline{K}\} \Longrightarrow \\
& \quad \exists i \in \{1,2,\dots,K\} \quad \ip{w_i}{x_t} > \ip{w_{\overline{y_t}}}{x_t} \; ?
\end{align*}
\end{framed}
\end{figure}

@@

\begin{proof}
$(\Longrightarrow)$ Let $S_1, S_2$ be the solution of the set splitting problem. Define
$$
w_1 = \left(a_1, a_2, \cdots, a_N, -\frac{1}{2}\right),
$$
where for all $i \in \{1,2,\dots,N\}$, $a_i=1$ if $i\in S_1$ and $a_i=-N$ if
$i\notin S_1$. Similarly, define
$$
w_2 = \left(b_1, b_2, \cdots, b_N, -\frac{1}{2}\right),
$$
where for all $i \in \{1,2,\dots,N\}$, $b_i=1$ if $i \in S_2$ and $b_i=-N$ if
$i\notin S_2$. Finally, define
$$
w_3 = (0,0,\cdots, 0),
$$
the zero vector. To see this is a solution for the weak labeling problem, we
verify separately for Type 1-3 samples defined above. For Type 1 sample, we have
$$
\ip{w_3}{x} = 0 > -\frac{1}{2} = \ip{w_1}{x}=\ip{w_2}{x}.
$$
For a Type 2 sample that corresponds to index $i$, we have either $i\in S_1$ or
$i\in S_2$ because $S_1\cup S_2 = \{1,2,\dots,N\}$ is guaranteed. Thus, either
$a_i=1$ or $b_i=1$. If $a_i=1$ is the case, then
$$
\ip{w_1}{x} = a_i - \frac{1}{2} = \frac{1}{2} > 0 = \ip{w_3}{x};
$$
similarly if $b_i=1$, we have $\ip{w_2}{x}>\ip{w_3}{x}$. \\ For a Type 3 sample
that corresponds to index $j$, Since $c_j \not\subset S_1$, there exists some
$i'\in c_j$ and $i'\notin S_1$. Thus we have $x_{i'}=1$, $a_{i'}=-N$, and
therefore
\begin{align*}
\ip{w_1}{x}
& = a_{i'}x_{i'} + \sum_{i\in \{1,2,\dots,N\} \setminus \{i'\}} a_ix_i - \frac{1}{2} \\
& \le -N + (N-1)-\frac{1}{2} \\
& < 0  \\
& = \ip{w_3}{x} \; .
\end{align*}
Because $c_j \not\subset S_2$ also holds, we also have
$\ip{w_2}{x}<\ip{w_3}{x}$. This direction is therefore proved. \\
\ \\
$(\Longleftarrow)$ Given the solution $w_1, w_2, w_3$ of the weak labeling problem, we define
\begin{align*}
S_1 &= \left\{i \in \{1,2,\dots,n\} ~:~ \ip{w_1-w_3}{(\e_i, 1)} > 0 \right\}, \\
S_2 &= \left\{i \in \{1,2,\dots,n\} ~:~ \ip{w_2-w_3}{(\e_i, 1)} > 0 \right. \\
& \qquad \qquad \qquad \left. \text{\ and\ } i \notin S_1 \right\}.
\end{align*}
It is not hard to see $S_1 \cap S_2 = \emptyset$ and $S_1\cup S_2 =
\{1,2,\dots,N\}$. The former is because $S_2$ only includes elements that are
not in $S_1$. For the latter, note that $(\e_i, 1)$ is the feature vector for
Type 2 samples. Because Type 2 samples all have label $\overline{3}$, for any $i
\in \{1,2,\dots,N\}$, one of the following must hold: $\ip{w_1-w_3}{(\e_i,
1)}>0$ or $\ip{w_2-w_3}{(\e_i, 1)}>0$. This implies $i\in S_1$ or $i\in S_2$.
